---
title: "Problem Set"
author: "Prof. Bisbee"
institute: "Yonsei University"
date: "Due Date: 2022/08/25 @ 9AM KST"
output:
  html_document: default
  pdf_document: default
---

## Getting Set Up

If you haven't already, create a folder for this course, and then two additional subfolders named `code` and `data`.

Open `RStudio` and create a new RMarkDown file (`.Rmd`) by going to `File -> New File -> R Markdown...`.
Change the title to `"Brain Korea 21: HW"` and the author to your full name. Save this file as `[LAST NAME]_HW.Rmd` to your `code` folder.

If you haven't already, download the `Pres2020_PV.Rds` file from the course [github page](https://github.com/jbisbee1/DSCI1000/blob/main/Lectures/Topic6_UnivariateVisualization/data/Pres2020_PV.Rds) and save it to your `data` folder.

All of the following questions should be answered using 
````{verbatim}
```{r}
# EXAMPLE CODE
mean(c(1,2,3))
```
````

## Question 1
Require `tidyverse` and load the `Pres2020_PV.Rds` data to `pres`.
```{r}
# INSERT CODE HERE
```

## Question 2
How many polls were fielded with sample sizes (`SampleSize`) larger than 2,000 respondents?
```{r}
# INSERT CODE HERE
```

> - Write answer here.

## Question 3
Which poll was "in the field" the longest? How many respondents did it have? When did it start and end?
```{r}
# INSERT CODE HERE
```

> - Write answer here.

## Question 4
Which polling company (`Poll`) conducted the most polls in the data?
```{r}
# INSERT CODE HERE
```

> - Write answer here.


## Question 5
Which poll was most favorable toward Biden? When was it fielded and how many respondents did it have? Answer the same question but for Trump. (NB: you will need to create a new variable with the `mutate()` command that measures the difference in support between Biden and Trump.)
```{r}
# INSERT CODE HERE
```

> - Write answer here.

## Question 6
How many polls predicted a Biden victory? How many predicted a Trump victory?
```{r}
# INSERT CODE HERE
```

> - Write answer here.

## Question 7
Using `ggplot`, visualize a histogram of sample sizes with `geom_histogram()`. Do the same but for the number of days in the field.
```{r}
# INSERT CODE HERE
```

## Question 8
Now plot a histogram of the difference between Biden support and Trump support. As above, you'll have to use the `mutate()` function to create this new variable. Now remake this plot but use `geom_bar()` instead of `geom_histogram()`.

```{r}
# INSERT CODE HERE
```

## Question 9
Create two barplots on the same graph, depicting Trump support (fill these bars with red) and Biden support (fill these bars with blue). Relabel the x and y axes to make sense to a reader who sees this graph for the first time.
```{r}
# INSERT CODE HERE
```

## Question 10
EXTRA CREDIT: Plot the total number of polls per start date in the data. NB: you will have convert `StartDate` to a `date` class with `as.Date()`. If you need help, see [this post](https://www.r-bloggers.com/2013/08/date-formats-in-r/).
```{r}
# INSERT CODE HERE
```
